Scaffolder - Software for \change{Reproducible}{Manual} Genome Scaffolding.

### 2011-08-14 src/article.tex
Revise discussion section

Scaffolding an incomplete genome assembly requires joining contigs and
additional gap-filling sequence using a combination of computational and
laboratory methods. The process of manually editing a scaffold is inherently
hard to reproduce and introduces the possibility of irreducible techniques
and/or human error. In respect to these drawbacks the aims of the Scaffolder
software are twofold: 1) to provide software that is easy to install and
simplifies the task of genome finishing; and 2) to facilitate reproducibility
in the scaffolding and finishing stage of a genome project. \pb

Scaffolder was designed to be as simple to use and, assuming the Ruby and
RubyGems software are present, can be installed in a single command line step.
This should negate any possible barriers to entry which require manual
compilation of source code. Scaffolder also uses a minimal and compact syntax
to describe how the genome scaffold sequence should be generated. This syntax
is simple to construct and edit whilst being succinct and readable. The goal
of this syntax is to make genome scaffolds easy to write in a common text
editor. \pb

Constructing a genome using by specifying the scaffold organisation in text
file makes generating a scaffold super sequence both reproducible and
deterministic for the same file and set of FASTA sequences. In comparison,
joining large nucleotide sequences by hand cannot be reliably reproduced. The
scaffold file furthermore provides a human readable description of how the
scaffold is constructed. Configuring the final sequence in the scaffold file
means the build is easier to edit once already constructed. The YAML text
format allows comparison of differences between scaffold builds using standard
Unix tools such as diff. This makes scaffold files easier to store in version
control systems and allows genome finishers to use methods similar to those in
software development. \pb

>>>>

Scaffolding an incomplete genome assembly requires joining contigs and
additional gap-filling sequence\add{s} using a combination of computational and
laboratory methods. The process of manually editing a scaffold is inherently
hard to reproduce and introduces \remove{the possibility of }irreducible
\change{techniques}{edits} and/or human error. In respect to \change{these
drawbacks}{this} the aims of the Scaffolder software are twofold: 1) to provide
software that is easy to install and simplifies the task of genome finishing;
and 2) to facilitate reproducibility in the scaffolding and finishing stage of
a genome project. \pb

\remove{Scaffolder was designed to be as simple to use and, assuming the Ruby
and RubyGems software are present, can be installed in a single command line
step. This should negate any possible barriers to entry which require manual
compilation of source code.} Scaffolder \remove{also} uses a minimal and
compact syntax to describe how \change{the}{a} genome scaffold sequence should
be generated. This syntax is simple to \change{construct}{write} and edit
whilst being succinct and readable. \change{The goal of this syntax is to make
genome scaffolds easy to write in a common text editor.}{The AGP format is
a similar format for describing scaffolds. This format can be used to define
contig order and N-filled gap regions in a scaffold. The advantages of the AGP
format is that each contig entry is defined on a single line which allows
searching the scaffold using UNIX command line tools. The scaffolder format in
comparison in contrast is written in the standardised YAML format and therefore
accessable to the many languages which provide parsers to this format. The
scaffolder format is provided with a tool explicitly to produce the fasta
sequence of scaffold specified by in the file.} \pb

Constructing a genome using by specifying the scaffold organisation in text
file makes generating a scaffold super sequence both reproducible and
deterministic for the same file and set of FASTA sequences. In comparison,
joining large nucleotide sequences by hand cannot be reliably reproduced. The
scaffold file furthermore provides a human readable description of how the
scaffold is constructed. Configuring the final sequence in the scaffold file
means the build is easier to edit once already constructed. \pb

\subsection*{Example Use Case}

\add{The example use case for scaffolder is a combination of compuation and
manual editing of a genome scaffold. As discussed many programming languages
provide libraries for reading and writing YAML. We have used scaffolder in our
genome projects create a scaffold file from parsing the output of \emph{in
silico} scaffolding tools into the scaffold file format. This scaffold was then
manually updated in a text editor with the results of using PCR to identify
gaps sequences. We see this as the advantage of describing a scaffold in both
a human and machine readable format.} \pb

The YAML text format \add{also} allows comparison of differences between
scaffold builds using standard Unix tools such as diff. This \add{therefore}
makes scaffold files \add{ameniable} to storage in version control systems and
allows genome finishers to use methods similar to those in software
development. \pb
